Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-02-05 article/02_Materials_and_Methods.Rmd
more consistent spacing before "calBC" (@yesdavid)
improving text in materials&methods

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
\caption{\label{map_graves}All graves dating between 2200 and 800calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}). The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape. The rectangular frame indicates the research area. The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from Natural Earth project (\url{www.naturalearthdata.com}).}
\end{figure}

The code of this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible. A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].

@@

The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013]. It contains ^14^C dates from the European Bronze and Early Iron Age (approximately 2300 BC to 500 BC). Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of more than 11000 published dates from around 2800 sites located all over Europe (as of December 2018). Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context. The variable *feature type* systematically distinguishes between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a.. This information allows to accurately and precisely determine a moment in space and time when a grave of a particular type was created. In theory, the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe. Unfortunately, the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no supervised or well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries. These issues though do not inhibit the general data analysis strategy of this paper.

>>>>

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../analysis/figures/map_graves.jpeg}
\caption{\label{map_graves}All graves dating between 2200 and 800 calBC as extracted from the Radon-B database (\url{radon-b.ufg.uni-kiel.de}). The classes of the variable \textit{burial type} are distinguished by colour, the ones of \textit{burial construction} by shape. The rectangular frame indicates the research area. The Map projection is EPSG:102013 and the base layer (land, rivers, lakes) data is taken from Natural Earth project (\url{www.naturalearthdata.com}).}
\end{figure}

The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018]. A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].

@@

The dataset used for the analysis in this paper is a subset of the openly available database Radon-B [@kneiselRadonB2013], which contains more than 11000 ^14^C dates from the European Bronze and Early Iron Age (approximately 2300 BC to 500 BC). Radiocarbon dating is a crucial absolute dating technique for this period [@ramseyRadiocarbonDatingRevolutions2008], and Radon-B is a collection of published dates from around 2800 sites located all over Europe (as of February 2019). Beyond the identifying laboratory number, the actual, uncalibrated date and a literature reference, the ^14^C dates in Radon-B are also documented with context information about the name and location of the dated site (coordinates), the sample material and the type and cultural attribution of the sampled context. The variable *feature type* systematically distinguishes between settlements, rock shelters and also different kinds of graves: *Grave (cremation), Grave (flat) inhumation*, i.a.. This information allows to accurately and precisely determine a moment in space and time when a grave of a particular type was created. In theory, the more radiocarbon dates with this information are available, the better becomes our understanding and reconstruction of burial traditions in prehistoric Europe. Unfortunately, the Radon-B dataset has some quality issues mostly due to inconsistent manual data input: Incompleteness and faultiness of meta information, no well-thought-out variable classification for *sample material* and *cultural attribution* and inconsistent multilingualism even within individual entries. These issues though do not inhibit the general data analysis strategy of this paper, as the information about temporal and spatial position as well as *feature type* attribution is mostly complete.

### 2019-02-05 article/05_Conclusion.Rmd
improving text in materials&methods

1. For this paper burial traditions were the main object of research, but the same methodological approach could be applied to reconstruct other aspects of cultural behaviour. The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives. Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances in anthropological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].

>>>>

1. For this paper burial traditions were the main object of research, but the same methodological approach could be applied to reconstruct other aspects of cultural behaviour. The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives. Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances in anthropological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017; @marwickPackagingData2018].
